---
output:
  html_document: default
---


```{r setup, include=FALSE}
# Load used packages
library(package = "readxl")
library(package = "tidyverse")

# configure ggplot2 theme
theme_set(
  theme_classic() +
    theme(
      title = element_text(face = "bold"),
      plot.subtitle = element_text(face = "plain")
    )
)

# configure knitting
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
knitr::opts_knit$set(root.dir = "..")
```


```{r data}
# load data
source(file = "scripts/load-data.R")
```


**Last updated:** `r substr(Sys.time(), 1, 19)`

# Statistical analysis

Statistical analysis performed using R v3.4.0 (R Core Team 2017).




# Lutzomyia diversity 2015

Based on data from 2015 collections, including:
site, trap, species and number of individuals.


## Data description

Follows some tables describing the data, for reference.

### Sex of collected specimens

```{r lutzomyia-sex-description}
lutzomyia_sex %>%
  mutate(
    sex = recode(
      sex,
      f = "female",
      m = "male"
    )
  ) %>%
  spread(key = sex, value = total_indivituals) %>%
  set_names(
    gsub("(^| )([a-z])", "\\1\\U\\2", gsub("_", " ", names(.)), perl = TRUE)
  ) %>%
  knitr::kable(
    caption = paste(
      "**Table 1:**",
      "Total individuals collected by species and sex" 
    )
  )
```


### Specimens by site

```{r lutzomyia-site-description}
lutzomyia %>%
  group_by(species, site) %>%
  summarize(
    n = n(),
    mean = mean(count),
    text = paste0(mean, " (", n, ")")
  ) %>%
  select(species, site, text) %>%
  spread(key = site, value = text) %>%
  set_names(
    gsub("(^| )([a-z])", "\\1\\U\\2", gsub("_", " ", names(.)), perl = TRUE)
  ) %>%
  set_names(
    ifelse(names(.) == "Species", names(.), paste0(names(.), ", mean (n)"))
  ) %>%
  knitr::kable(
    caption = paste(
      "**Table 2:**",
      "Mean number of collected individuals by species and forest fragmentation"
    )
  )
```


## Results

Follow results for the statistical analysis.

### Diversity

```{r prepare-diversity-data}
#------------------------------------------------------------------------------*
# ldutari:
# Indices de diversidad y curvas de acumulacion
# (los indices pueden ser Alpha de Fisher, Margalef, Shannon, Simpson,
# Berger-Parker)
#------------------------------------------------------------------------------*

# All trap order permutations
trap_permutations <- lutzomyia %>%
  # For each site and its traps
  select(site, trap) %>%
  split(.$site) %>%
  map(unique) %>%
  # Calculate all permutations of traps (i.e. simulate order of use)
  map(~combinat::permn(.x$trap)) %>%
  # Put in nice rectangular structure
  map(simplify2array) %>%
  map(t) %>%
  map(as_data_frame) %>%
  # Combine everything
  bind_rows(.id = "site") %>%
  group_by(site) %>%
  mutate(permutation = 1:n()) %>%
  # Use tidy structure
  set_names(gsub("V", "", names(.))) %>%
  gather(key = effort, value = trap, convert = TRUE, -site, -permutation) %>%
  arrange(site, permutation)
```



```{r cummulative-species}
# Number of species collected per trap per site
collected_species <- lutzomyia %>%
  # Number of species by trap
  group_by(site, trap) %>%
  summarize(
    species = sum(count > 0)
  )

# Cummulative curve using traps as effort
cum_curve_traps <- collected_species %>%
  # Bind with trap permutations
  right_join(trap_permutations) %>%
  # Cummulative species by number of traps
  group_by(site, permutation) %>%
  mutate(
    cum_species = cumsum(species)
  ) %>%
  # Mean number of cummulative species by effort (effort = number of traps)
  group_by(site, effort) %>%
  summarize(
    species_sd = sd(cum_species),
    species_mean = mean(cum_species)
  ) %>%
  ungroup
```


#### Heterogeneity

```{r aov-heterogeneity}
# aov model
sites_aov <- collected_species %>%
    aov(
        formula = species ~ site
    )

print(sites_aov)
print(summary(sites_aov))
```


A one-way analysis of variance (ANOVA) was used to test for heterogeneity in
the number of species collected per trap between the sites,
and found to be not significantly heterogeneous
(one-way anova,
F~`r paste(summary(sites_aov)[[1]]$Df, collapse = ",")`~ =
`r signif(na.omit(summary(sites_aov)[[1]][,"F value"]), 2)`,
p = `r signif(na.omit(summary(sites_aov)[[1]][,"Pr(>F)"]), 2)`).



#### Cummulative species curve

**Figure 1** shows the cummulative species curve per site given the number of
traps (Jimemez 2003).
The curve for each site was calculated by averaging the number of detected
species accumulated when adding traps,
over all the possible permutations for the order in which each one of the 5
traps is added to the calculation.


```{r cummulative-curve-traps}
# Plot cummulative curve
cum_curve_traps %>%
  mutate(
    site = gsub("(^| )([a-z])", "\\1\\U\\2", gsub("_", " ", site), perl = TRUE),
    p1sd = species_mean + species_sd,
    m1sd = species_mean - species_sd
  ) %>%
  ggplot(aes(x = effort, y = species_mean, color = site)) +
  geom_line() +
  # geom_line(aes(y = p1sd), linetype = "dashed") +
  # geom_line(aes(y = m1sd), linetype = "dashed") +
  labs(
    title = paste(
      "Figure 1: Cummulative species curve per site",
      "by number of traps.",
      sep = "\n"
    ),
    subtitle = paste(
      "Number of species calculated by averaging the cummulative species",
      "over all permutations of the traps in each site.",
      sep = "\n"
    ),
    x = "Number of traps",
    y = "Number of observed species",
    color = "Site"
  ) +
  coord_fixed(ratio = 1/10)
```



# Lutzomyia diversity 2010-2014

## Data description


## Results

### Species relative abundance between canopy and underbrush

> Pending vertical strata data

```{r vertical-relative-abundance}
#------------------------------------------------------------------------------*
# ldutari:
# Abundancia relative de especies mas abundantes entre dosel y el sotobosque,
# utilizando T-test y U de Mann- Whitney
#------------------------------------------------------------------------------*
```



# References

Jimenez, A (2003). Las curvas de acumulación de especies y la necesidad de
evaluar la calidad de los inventarios biológicos. Revista Ibérica de Aracnología.
Vol. 8, 31-XII-2003, 151-161.

R Core Team (2017). R: A language and environment for statistical computing.
R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.
